Bioverse: Potentially Observable Exoplanet Biosignature Patterns Under the UV Threshold Hypothesis for the Origin of Life

### 2025-02-04 src/tex/ms.tex
added some discussion in 4.1 addressing Reviewers 1 comment number 2

Presumably, not all habitable worlds are inhabited and not all inhabited worlds develop detectable biosignatures.
The fraction of \glspl{EEC} that are both inhabited and exhibit detectable biosignatures at the time of observation is unknown \rev{and is represented by the term $f_\mathrm{life}$.
This encompasses the probability of life both emerging and persisting to produce detectable biosignatures.}
Let us consider the probability to detect a biosignature $P(L)$, and let our observable be the inferred past \gls{NUV} flux of the planet $F_\mathrm{NUV}$.
Under Hypothesis $H_1$, there exists a special unknown value of $F_\mathrm{NUV}$, noted $F_\mathrm{NUV, min}$ such that

@@

\subsection{Semi-analytical assessment}\label{sec:results-semianalytical}
In Section~\ref{sec:met-semianalytical}, we computed the probability for true positive evidence for $H_\mathrm{1}$ and $H_\mathrm{null}$, respectively (Equations~\ref{eq:bayes_factor},~\ref{eq:bayes_factor2}).
Figure~\ref{fig:semian_true_evidence} shows how these evidences are distributed for sample sizes \var{semian_Nsamp1} and \var{semian_Nsamp2}, and how likely we are to obtain strong evidence ($BF_{H_i, H_j}$ > 10).
For $n = \var{semian_Nsamp1}$, strong true evidence for $H_\mathrm{1}$ ($H_\mathrm{null}$) can be expected in $\sim \SI{30}{\percent}$ ($\sim \SI{40}{\percent}$) of all random experiments.
In the majority of cases, the outcome of the survey will be inconclusive.
\begin{figure*}
    \script{semian_true_evidence.py}
    \begin{centering}
        \includegraphics[width=\hsize]{figures/semian_true_evidence}
        \caption{Obtaining true strong evidence with different sample sizes. Left: Probability to reach given evidence levels for $H_\mathrm{1}$ and $H_\mathrm{null}$ under sample sizes $n = \var{semian_Nsamp1}$ (solid) and $n = \var{semian_Nsamp2}$ (dashed). Vertical lines denote thresholds for ``strong'' evidence, $BF_{H_i, H_j}$ > 10, and ``extreme'' evidence, $BF_{H_i, H_j}$ > 100. Right: Probability of obtaining true strong evidence for $H_\mathrm{1}$ as a function of sample size $n$.}
        \label{fig:semian_true_evidence}
    \end{centering}
\end{figure*}
The situation improves with larger samples: for $n = \var{semian_Nsamp2}$, \SI{80}{\percent} of random samples permit conclusive inference (strong true evidence) under either $H_\mathrm{1}$ or $H_\mathrm{null}$.

@@

So far, we have assumed random, uniform distributions of $f_\mathrm{life}$, $F_\mathrm{NUV, min}$, and $F_\mathrm{NUV}$. % Actually it is more than just NUV, it is whatever index that integrates NUV and HZ requiremetns into a single scalar variable for which a threshold exist
A high biosignature detection rate $f_\mathrm{life}$ increases the evidence (see Equation~\ref{eq:bayes_factor}) but a survey strategy cannot influence it.
The same is true for $F_\mathrm{NUV, min}$, where again higher values increase the evidence as the binomial distribution for $H_\mathrm{1}$ gets increasingly skewed and shifted away from the one for $H_\mathrm{null}$.
However, one might select exoplanets for which a biosignature test is performed based on \textit{a priori} available contextual information~\citep{catling2018exoplanet} in order to maximize the science yield of investing additional resources.
For instance, the distribution of $F_\mathrm{NUV}$ in the planet sample can be influenced by the survey strategy, and a targeted sampling approach could favor extreme values. % of $F_\mathrm{NUV}$ in the sample selection.
We model this by distributing $F_\mathrm{NUV}$ according to different $Beta$ functions and introduce a selectivity parameter $s\in]-1,1[$ such that $F_\mathrm{NUV} \sim Beta(1/10^s,1/10^s)$.
Figure~\ref{fig:semian_selectivity} shows how the probability of obtaining true strong evidence for $H_\mathrm{1}$ scales with selectivity $s$.
%Here, $s=0$ corresponds to a random uniform distribution.
%Compared to this case, a high selectivity ($s \sim 1$) can increase the probability of obtaining true strong evidence to $\gtrsim \SI{85}{\percent}$ for large samples.
For large samples, a high selectivity ($s \sim 1$) can increase the probability of obtaining true strong evidence from $\sim \SI{70}{\percent}$ for $s= 0$ (random uniform distribution) to $> \SI{90}{\percent}$.

\begin{figure*}
    \script{semian_true_evidence.py}
    \begin{centering}
        \includegraphics[width=\hsize]{figures/semian_selectivity}
        \caption{Scaling of the probability of obtaining true strong evidence with sample selectivity. Left: Sampling distribution for different selectivity parameters $s$. Right: Resulting \mbox{P(true strong evidence)}, where $f_\mathrm{life}$ and $F_\mathrm{NUV, min}$ are randomly drawn from a uniform distribution. Sampling more extreme values of $F_\mathrm{NUV}$ is more likely to yield strong evidence.}
        \label{fig:semian_selectivity}
    \end{centering}
\end{figure*}

@@

>>>>

Presumably, not all habitable worlds are inhabited and not all inhabited worlds develop detectable biosignatures.
The fraction of \glspl{EEC} that are both inhabited and exhibit detectable biosignatures at the time of observation is unknown \rev{and is represented by the term $f_\mathrm{life}$.
This encompasses the probability of life both emerging and persisting to produce detectable biosignatures.}
\rev{Due to our ignorance about its true value or even its order of magnitude, we draw $f_\mathrm{life}$ from a log-uniform prior probability distribution.}
Let us consider the probability to detect a biosignature $P(L)$, and let our observable be the inferred past \gls{NUV} flux of the planet $F_\mathrm{NUV}$.
Under Hypothesis $H_1$, there exists a special unknown value of $F_\mathrm{NUV}$, noted $F_\mathrm{NUV, min}$ such that

@@

\subsection{Semi-analytical assessment}\label{sec:results-semianalytical}
In Section~\ref{sec:met-semianalytical}, we computed the probability for true positive evidence for $H_\mathrm{1}$ and $H_\mathrm{null}$, respectively (Equations~\ref{eq:bayes_factor},~\ref{eq:bayes_factor2}).
Figure~\ref{fig:semian_true_evidence} shows how these evidences are distributed for sample sizes \var{semian_Nsamp1} and \var{semian_Nsamp2}, and how likely we are to obtain strong evidence ($BF_{H_i, H_j}$ > 10) \rev{in the agnostic case where we draw $f_\mathrm{life}$ from a log-uniform distribution}.
For $n = \var{semian_Nsamp1}$, strong true evidence for $H_\mathrm{1}$ ($H_\mathrm{null}$) can be \rev{expected in $\sim \SI{3}{\percent}$ ($\sim \SI{6}{\percent}$) of} all random experiments.
In the majority of cases, the outcome of the survey will be inconclusive.
\begin{figure*}
    \script{semian_true_evidence.py}
    \begin{centering}
        \includegraphics[width=\hsize]{figures/semian_true_evidence}
        \caption{Obtaining true strong evidence with different sample sizes. Left: Probability to reach given evidence levels for $H_\mathrm{1}$ and $H_\mathrm{null}$ under sample sizes $n = \var{semian_Nsamp1}$ (solid) and $n = \var{semian_Nsamp2}$ (dashed). Vertical lines denote thresholds for ``strong'' evidence, $BF_{H_i, H_j}$ > 10, and ``extreme'' evidence, $BF_{H_i, H_j}$ > 100. Right: Probability of obtaining true strong evidence for $H_\mathrm{1}$ as a function of sample size $n$.}
        \label{fig:semian_true_evidence}
    \end{centering}
\end{figure*}
The situation improves with larger samples: \rev{for $n = \var{semian_Nsamp2}$, \SI{14}{\percent} (\SI{16}{\percent}) of} random samples permit conclusive inference (strong true evidence) \rev{under $H_\mathrm{1}$ ($H_\mathrm{null}$).}

@@

So far, we have \rev{drawn random values from uniform distributions for} $F_\mathrm{NUV, min}$, and $F_\mathrm{NUV}$\rev{, and from a log-uniform distribution for $f_\mathrm{life}$}. % Actually it is more than just NUV, it is whatever index that integrates NUV and HZ requiremetns into a single scalar variable for which a threshold exist
A high biosignature detection rate $f_\mathrm{life}$ increases the evidence (see Equation~\ref{eq:bayes_factor}) but a survey strategy cannot influence it.
The same is true for $F_\mathrm{NUV, min}$, where again higher values increase the evidence as the binomial distribution for $H_\mathrm{1}$ gets increasingly skewed and shifted away from the one for $H_\mathrm{null}$.
However, one might select exoplanets for which a biosignature test is performed based on \textit{a priori} available contextual information~\citep{catling2018exoplanet} in order to maximize the science yield of investing additional resources.
For instance, the distribution of $F_\mathrm{NUV}$ in the planet sample can be influenced by the survey strategy, and a targeted sampling approach could favor extreme values. % of $F_\mathrm{NUV}$ in the sample selection.
We model this by distributing $F_\mathrm{NUV}$ according to different $Beta$ functions and introduce a selectivity parameter $s\in]-1,1[$ such that $F_\mathrm{NUV} \sim Beta(1/10^s,1/10^s)$.
Figure~\ref{fig:semian_selectivity} shows how the probability of obtaining true strong evidence for $H_\mathrm{1}$ scales with selectivity $s$.
%Here, $s=0$ corresponds to a random uniform distribution.
%Compared to this case, a high selectivity ($s \sim 1$) can increase the probability of obtaining true strong evidence to $\gtrsim \SI{85}{\percent}$ for large samples.
For large samples, a high selectivity ($s \sim 1$) can increase the probability of obtaining true strong evidence from $\sim \SI{70}{\percent}$ for $s= 0$ (random uniform distribution) to $> \SI{90}{\percent}$.

\begin{figure*}
    \script{semian_true_evidence.py}
    \begin{centering}
        \includegraphics[width=\hsize]{figures/semian_selectivity}
        \caption{Scaling of the probability of obtaining true strong evidence with sample selectivity. Left: Sampling distribution for different selectivity parameters $s$. Right: Resulting \mbox{P(true strong evidence)}, where $f_\mathrm{life}$ and $F_\mathrm{NUV, min}$ are randomly drawn from \rev{log-uniform and uniform distributions, respectively}. Sampling more extreme values of $F_\mathrm{NUV}$ is more likely to yield strong evidence.}
        \label{fig:semian_selectivity}
    \end{centering}
\end{figure*}

@@

\appendix
\section{Alternative priors for the abiogenesis rate}
\label{sec:appendix-flife-prior}
\rev{In the main text, we used a log-uniform prior for the probability of life emerging and persisting $f_\mathrm{life}$ in order to reflect our ignorance about this parameter and its order of magnitude.
We consider this a reasonable choice, as it is agnostic about the scale of $f_\mathrm{life}$ and assigns equal prior probability to all orders of magnitude.
Nevertheless, we repeated our semi-analytical analysis for the more optimistic choice of a prior that is \textit{uniform} in $f_\mathrm{life}$.
Figure~\ref{fig:semian_true_evidence_uniform} shows the resulting probabilities of obtaining strong evidence for or against the UV~Threshold Hypothesis under this prior.
While the resulting trends are qualitatively similar, the uniform prior yields overall higher probabilities of such a conclusive test for all sample sizes.
}

\begin{figure*}
    \begin{centering}
        \includegraphics[width=\hsize]{figures/semian_true_evidence_uniform}
        \caption{\rev{As Figure~\ref{fig:semian_true_evidence}, but using a uniform prior for the abiogenesis rate $f_\mathrm{life}$.
        The probabilities of obtaining strong evidence for or against the UV~Threshold Hypothesis are overall higher than for the log-uniform prior.
        }
        }
        \label{fig:semian_true_evidence_uniform}
    \end{centering}
\end{figure*}
